Robustness of MinD oscillation in \emph{Escherichia coli} with diverse cell shapes

### 2014-02-26 paper/paper.tex
wrote the > shape section and updated one of the data files for it Tested everything on node22.

\begin{figure}
  \includegraphics[width=4.2cm]{../data/shape-randst/plots/box-plot_D--randst-25-800-800-9500-1500}
  \includegraphics[width=4.2cm]{../data/shape-randst/plots/box-plot_D--randst-25-1000-1700-9400-1500}
  \caption{Box plots of Mannik shapes for sizes at the lower stability
    threshhold.  The vertical axis shows stacked and changing over
    time the total number of proteins that are of four different
    compound stages and in the different sections of the cell.  Both
    shapes have longer axis of roughly 2.2$\mu m$.}
  \label{box-mannik}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=2.8cm]{../mannik-2.png}
  \includegraphics[width=2.8cm]{../mannik-1.png}
  \includegraphics[width=\columnwidth]{../data/shape-randst/plots/paper-arrow-plot}
  \caption{Arrow plots for our simulation and mannik's experimental}
  \label{arrow-compare-mannik}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=4.2cm]{../data/shape-triangle/plots/box-plot_D--triangle-25-233-233-233-1500}
  \includegraphics[width=4.2cm]{../data/shape-triangle/plots/box-plot_D--triangle-25-266-266-266-1500}
  %\includegraphics[width=4.2cm]{../data/shape-triangle/plots/box-plot_D--triangle-25-800-800-800-1500}
  \caption{Two box plots of triangle shapes.  The two smaller show
    sizes at the lower stability threshhold, one just below, one just
    above.  The larger triangle is the largest simulated. The vertical
    axis shows stacked and changing over time the total number of
    proteins that are of four different compound stages and in the
    different sections of the cell.}
  \label{box-triangle}
\end{figure}

@@

\section{Our '\textgreater' Shaped Cells}
\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-525-525-9900-1500}
  \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-550-550-9900-1500}
  \label{>-box}
\end{figure}
\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-600-600-9900-1500}
  \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-1400-1400-9900-1500}
  \label{>-box2}
\end{figure}

We have simulated a flattened '>' shape that shows an interesting size
dependent pattern (seen in Figures~\ref{>-box}~and~\ref{>-box2}).
When the shape is small enough that the distance from the apex tip to
the opposite curve is less than our stability limit, a simple, two
pole oscillation pattern develops and is sustained (where the two
poles are on opposite leftward corners of the cell).  When the cell is
increased in size, so that this distance measures approximately equal
to our stability value, the cell develops a double oscillation in
time, where the maxima oscillates from the lower corner to the
rightward apex, then \emph{back to the lower corner} then to the apex
again, then to the upper corner, then back to the apex, then to the
upper corner, then to the apex again, then back to the lower corner,
and continues this pattern indefinately.  Figure~\ref{>-box} shows the
smaller size, which oscillates simply, and an intermediate size, which
shows a pattern for which one out of three spikes in the regular
oscillation is less significant.  This second pattern seems to be a
precursor to the double oscillation seen in Figure~\ref{>-box2}.  The
second plot shown in Figure~\ref{>-box2} shows a slighlty larger cell,
for which the distance measured is larger than our stability limit
distance.  This shows a very random pattern.

>>>>

\begin{figure}
%  \includegraphics[width=4.2cm]{../data/shape-randst/plots/box-plot_D--randst-25-800-800-9500-1500}
%  \includegraphics[width=4.2cm]{../data/shape-randst/plots/box-plot_D--randst-25-1000-1700-9400-1500}
  \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-800-800-9500-1500}
  \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-1000-1700-9400-1500}
  \caption{Box plots of Mannik shapes for sizes at the lower stability
    threshhold.  The vertical axis shows stacked and changing over
    time the total number of proteins that are of four different
    compound stages and in the different sections of the cell.  Both
    shapes have longer axis of roughly 2.2$\mu m$.}
  \label{box-mannik}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-randst/plots/paper-arrow-plot}
  \caption{Arrow plots for our simulation and mannik's experimental}
  \label{arrow-compare-mannik}
\end{figure}

@@

\begin{figure}
%  \includegraphics[width=4.2cm]{../data/shape-triangle/plots/box-plot_D--triangle-25-233-233-233-1500}
%  \includegraphics[width=4.2cm]{../data/shape-triangle/plots/box-plot_D--triangle-25-266-266-266-1500}
  \includegraphics[width=\columnwidth]{../data/shape-triangle/plots/box-plot_D--triangle-25-233-233-233-1500}
  \includegraphics[width=\columnwidth]{../data/shape-triangle/plots/box-plot_D--triangle-25-266-266-266-1500}
  %\includegraphics[width=4.2cm]{../data/shape-triangle/plots/box-plot_D--triangle-25-800-800-800-1500}
  \caption{Two box plots of triangle shapes.  The two smaller show
    sizes at the lower stability threshhold, one just below, one just
    above.  The larger triangle is the largest simulated. The vertical
    axis shows stacked and changing over time the total number of
    proteins that are of four different compound stages and in the
    different sections of the cell.}
  \label{box-triangle}
\end{figure}

@@

\section{Our '\textgreater' Shaped Cells}
\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-525-525-9900-1500}
  \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-550-550-9900-1500}
  \label{Vbox}
\end{figure}
\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-600-600-9900-1500}
  \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-1400-1400-9900-1500}
  \label{Vbox2}
\end{figure}

We have also simulated a flattened '\textgreater' shape that shows an
interesting size dependent pattern (seen in
Figures~\ref{Vbox}~and~\ref{Vbox2}).  When the shape is small enough
that the distance from the apex tip to the opposite curve is less than
our stability limit, a simple, two pole oscillation pattern develops
and is sustained (where the two poles are on opposite leftward corners
of the cell).  When the cell is increased in size so that this
distance is approximately equal to our stability value, the cell
develops a double oscillation in time, where the maxima oscillates
from the lower corner to the rightward apex, then \emph{back to the
  lower corner}, then to the apex again, then to the upper corner,
then back to the apex, then to the upper corner, then to the apex
again, then back to the lower corner, and continues this pattern
indefinately.  Figure~\ref{Vbox} shows the smaller size, for which the
protein oscillates simply, and an intermediate size, which shows a
pattern for which one out of three spikes in the regular oscillation
is less significant.  This second pattern seems to be a precursor to
the double oscillation that seen in Figure~\ref{Vbox2}.  The second
plot shown in Figure~\ref{Vbox2} shows a slightly larger cell for
which the distance measured is larger than our stability limit
distance.  This simulation shows a random pattern.

A detailed analysis of a time series of protein contour plots shows
that the importance of spatial dimensions is based on the ability of
protein in different stages to diffuse neccesary distances before
undergowing certain types of reactions.  The double oscillation
pattern shows that every time the membrane bound MinD maximizes in the
upper corner as opposed to the rightward apex, it does so because the
MinE has not had enough time to diffuse to this upper corner, leaving
a MinE free zone.  This particular pattern is specific to this shape
and size, but in general its the restrictions set in place by the
rates at which proteins diffuse and the distances they cover before
reacting that dictate their oscillatory bahavoir.

Simulations of this shape in which the two planar dimensions are large
compared to those shown(roughly $9\micron$ across) show no stable
regular periodic pattern. Animations show that the MinD acts as a sort
of traveling cluster that is constantly being eaten away by MinE
behind it.  When this cluster enters a region of high curvature that
ends in a dead end corner it performs the polar maxima behavoir seen
in the other simulations, regrouping out away from the corner where it
travels in the opposite direction while being ``chased'' away by the
MinE that itself is currently maximizing in the corner.  The maxima
occur on non-polar walls often, and will often split into two roaming
clusters.
